path-independence

### 2020-02-25 main.tex

The naive approach to verifying path independence would be to compare all conflicting paths pairwise and thus ensure that there are no instances of violation of the property. Over the course of online addition of an edge, given that the graph before adding the edge was path independent, it would only be necessary to check conflicting pairs involving a new path that were created as a consequence of adding the new edge to the path.
To identify all such conflicting pairs, we present the \textit{two-flip tolerant path search}.

@@

\paragraph{Theorem} The result of the two-flip tolerant path search from the source to the sink node of an edge that is to be newly added has a one-to-one* correspondence to the set of new pairs of conflicting paths that arise in the graph due to its addition.

*= This excludes paths whose "area of difference" does not include the new edge since they should remain path independent.

\paragraph{Proof}
Let the new edge added to the graph be (S, T).
It has already been shown how each path corresponds to a conflicting pair in the case where a path has two flips. When a path has only the first flip (which is to say, the third phase of the path is missing), but not the second flip, sink node T can be treated as the sink of the two conflicting path, while the first flipping point remains the source. Similarly when only the second flipping point is present then it is the sink, and S is the source. Finally when no flipping points are present, there are two possibilities. 
Either the path is oriented from S to T, in which case the conflicting paths are simply the edge (S, T) and the entire flip-less path.
Or, the path is oriented from T to S. In this case, the conflicting paths are the identity on S (that implicitly exists) and the cycle obtained from composing the new edge with the path.

@@

\subsection{Verification using the new paths}
The flip tolerant paths search returns pairs of conflicting paths. Verification is performed by composing edge functions to derive path functions, and then performing an equality check to verify if the path functions are equal.

@@

% \subsubsection{Theorem: If two paths have one flip-point in common and there exist some path connecting the second flip points, then there exists a partial ordering between the two paths}

% Two possible cases:

% \begin{center}
% \begin{tikzpicture}[scale=0.2]
% \tikzstyle{every node}+=[inner sep=0pt]
% \draw [black] (25,-31.9) circle (3);
% \draw (25,-31.9) node {$S$};
% \draw [black] (48,-31.9) circle (3);
% \draw (48,-31.9) node {$T$};
% \draw [black] (26.6,-13.6) circle (3);
% \draw (26.6,-13.6) node {$F_1$};
% \draw [black] (39,-21) circle (3);
% \draw (39,-21) node {$F_2$};
% \draw [black] (45.1,-12.9) circle (3);
% \draw (45.1,-12.9) node {$F_3$};
% \draw [black] (29.18,-15.14) -- (36.42,-19.46);
% \fill [black] (36.42,-19.46) -- (35.99,-18.62) -- (35.48,-19.48);
% \draw (31.38,-17.8) node [below] {$g_2$};
% \draw [black] (23.529,-29.292) arc (-156.64712:-213.3464:14.141);
% \fill [black] (23.53,-29.29) -- (23.67,-28.36) -- (22.75,-28.76);
% \draw (21.8,-22.39) node [left] {$f_1$};
% \draw [black] (27.889,-16.304) arc (19.98764:-29.98116:15.634);
% \fill [black] (26.74,-29.46) -- (27.57,-29.02) -- (26.71,-28.52);
% \draw (29.4,-23.08) node [right] {$f_2$};
% \draw [black] (46.09,-29.59) -- (40.91,-23.31);
% \fill [black] (40.91,-23.31) -- (41.03,-24.25) -- (41.81,-23.61);
% \draw (42.95,-27.88) node [left] {$h_2$};
% \draw [black] (29.6,-13.49) -- (42.1,-13.01);
% \fill [black] (42.1,-13.01) -- (41.28,-12.54) -- (41.32,-13.54);
% \draw (35.81,-12.7) node [above] {$g_1$};
% \draw [black] (47.55,-28.93) -- (45.55,-15.87);
% \fill [black] (45.55,-15.87) -- (45.18,-16.73) -- (46.17,-16.58);
% \draw (47.25,-22.21) node [right] {$h_1$};
% \draw [black] (40.8,-18.6) -- (43.3,-15.3);
% \fill [black] (43.3,-15.3) -- (42.41,-15.63) -- (43.21,-16.24);
% \draw (42.63,-18.35) node [right] {$l$};
% \end{tikzpicture}
% \end{center}

% $P_2$ is better, intuitively because l can only lose- not add information.

% \begin{center}
% \begin{tikzpicture}[scale=0.2]
% \tikzstyle{every node}+=[inner sep=0pt]
% \draw [black] (25,-31.9) circle (3);
% \draw (25,-31.9) node {$S$};
% \draw [black] (48,-31.9) circle (3);
% \draw (48,-31.9) node {$T$};
% \draw [black] (26.7,-7.7) circle (3);
% \draw (26.7,-7.7) node {$F_1$};
% \draw [black] (43.1,-21.8) circle (3);
% \draw (43.1,-21.8) node {$F_2$};
% \draw [black] (29.9,-16.8) circle (3);
% \draw (29.9,-16.8) node {$F_3$};
% \draw [black] (29.408,-6.43) arc (107.51449:-8.88948:11.282);
% \fill [black] (43.95,-18.93) -- (44.57,-18.22) -- (43.58,-18.06);
% \draw (41.59,-8.14) node [above] {$g_1$};
% \draw [black] (22.914,-29.751) arc (-141.57162:-226.46499:15.013);
% \fill [black] (22.91,-29.75) -- (22.81,-28.81) -- (22.02,-29.44);
% \draw (19.1,-19.32) node [left] {$f_1$};
% \draw [black] (46.06,-22.008) arc (71.20036:-19.43982:5.814);
% \fill [black] (46.06,-22.01) -- (46.66,-22.74) -- (46.98,-21.79);
% \draw (50.12,-23.9) node [right] {$h_2$};
% \draw [black] (45.319,-30.601) arc (-127.57654:-180.66292:7.327);
% \fill [black] (42.46,-24.71) -- (41.95,-25.5) -- (42.95,-25.52);
% \draw (42.49,-29.07) node [left] {$h_1$};
% \draw [black] (28.97,-19.65) -- (25.93,-29.05);
% \fill [black] (25.93,-29.05) -- (26.65,-28.44) -- (25.7,-28.13);
% \draw (26.68,-23.67) node [left] {$f_2$};
% \draw [black] (32.71,-17.86) -- (40.29,-20.74);
% \fill [black] (40.29,-20.74) -- (39.72,-19.99) -- (39.37,-20.92);
% \draw (35.16,-19.83) node [below] {$g_2$};
% \draw [black] (27.7,-10.53) -- (28.9,-13.97);
% \fill [black] (28.9,-13.97) -- (29.11,-13.05) -- (28.17,-13.38);
% \end{tikzpicture}
% \end{center}

% $P_1$ is better for a similar reason.

\section{Optimization}

@@

\subsection{Polynomial verification bound}
 This is a special instance of case 1, where l and m are the identity.
\begin{center}
\begin{tikzpicture}[scale=0.2]
\tikzstyle{every node}+=[inner sep=0pt]
\draw [black] (25,-31.9) circle (3);
\draw (25,-31.9) node {$S$};
\draw [black] (48,-31.9) circle (3);
\draw (48,-31.9) node {$T$};
\draw [black] (28.9,-19.5) circle (3);
\draw (28.9,-19.5) node {$F_1$};
\draw [black] (43.9,-19.5) circle (3);
\draw (43.9,-19.5) node {$F_2$};
\draw [black] (30.687,-17.11) arc (132.96162:47.03838:8.383);
\fill [black] (42.11,-17.11) -- (41.87,-16.2) -- (41.19,-16.93);
\draw (36.4,-14.36) node [above] {$g_1$};
\draw [black] (41.409,-21.156) arc (-63.93383:-116.06617:11.399);
\fill [black] (41.41,-21.16) -- (40.47,-21.06) -- (40.91,-21.96);
\draw (36.4,-22.82) node [below] {$g_2$};
\draw [black] (23.657,-29.237) arc (-164.08097:-230.83744:7.929);
\fill [black] (23.66,-29.24) -- (23.92,-28.33) -- (22.96,-28.61);
\draw (22.95,-24.04) node [left] {$f_1$};
\draw [black] (30.007,-22.272) arc (11.75211:-46.67052:8.58);
\fill [black] (27.49,-30.26) -- (28.42,-30.08) -- (27.73,-29.35);
\draw (30.56,-27.24) node [right] {$f_2$};
\draw [black] (45.566,-30.169) arc (-134.8387:-188.56886:9.123);
\fill [black] (42.98,-22.34) -- (42.36,-23.06) -- (43.35,-23.21);
\draw (42.57,-27.26) node [left] {$h_2$};
\draw [black] (46.175,-21.439) arc (41.169:-4.57656:10.227);
\fill [black] (46.18,-21.44) -- (46.33,-22.37) -- (47.08,-21.71);
\draw (48.95,-24.27) node [right] {$h_1$};
\end{tikzpicture}
\end{center}

@@

In an N node graph, there are $^NC_2 = \frac{N(N-1)}{2}$ choices of flip points. Thus we can place a polynomial upper bound on the number of paths that actually need to be verified.

@@

\subsection{Optimized Path Finding Algorithm}
We utilize this idea of selecting flip points to devise a polynomial time path search. This search does not return all two flip tolerant paths, but rather, a path for every feasible choice of flipping point.
 *TODO Write algorithm.*

>>>>

The naive approach to verifying path independence would be to compare all conflicting paths pairwise and thus ensure that there are no instances of violation of the property. Over the course of online addition of an edge, given that the graph before adding the edge was path independent, it would only be necessary to check conflicting pairs involving a new path that were created as a consequence of adding the new edge to the path.
To identify all such conflicting pairs, we present the \textit{two-flip tolerant path search}. This naive approach may be considered a baseline solution.

@@

The \textit{path extraction algorithm} then transforms the output of the two flip path search to the set of new pairs of conflicting paths that arise in the graph due to its addition.

*= This excludes paths whose "area of difference" does not include the new edge since they should remain path independent. It also excludes paths that in their course contain cycles that exclude the new edge, since these paths must, again by path independence of the original graph, be equal to the path resulting from removing their cycles.

\paragraph{Path extraction algorithm}
The purpose of this algorithm is to derive the conflicting pairs that a given two-flip tolerant path corresponds to.

Let the new edge added to the graph be (S, T).
It has already been shown how each path corresponds to a conflicting pair in the case where a path has two flips. When a path has only the first flip (which is to say, the third phase of the path is missing), but not the second flip, sink node T can be treated as the sink of the two conflicting path, while the first flipping point remains the source. Similarly when only the second flipping point is present then it is the sink, and S is the source. Finally when no flipping points are present, there are two possibilities. 
Either the path is oriented from S to T, in which case the conflicting paths are simply the edge (S, T) and the entire flip-less path.
Or, the path is oriented from T to S. In this case, the conflicting paths are the identity on S (that implicitly exists) and the cycle obtained from composing the new edge with the path. The cycle conflicts with all other cycles starting from each node in the cycle. Notice that a single cycle can give rise to multiple conflicting pairs.

\paragraph{Theorem} The result of the two-flip tolerant path search from the source to the sink node of an edge that is to be newly added, followed with the path extraction algorithm has a one-to-one* correspondence to the set of new pairs of conflicting paths that arise in the graph due to its addition.

*= Same conditions as above.

\paragraph{Proof}

Every element in the output of the path extraction algorithm was by definition a conflicting pair.

@@

*Insert python code here*

\subsection{Implementation of path extraction}
 *Again, insert code*

\subsection{Verification using the new paths}
The algorithm returns pairs of conflicting paths. Verification is performed by composing edge functions to derive path functions, and then performing an equality check to verify if the path functions are equal.

@@

\subsection{Optimization of two-flip search}

@@

\section{Optimization}
We now present a polynomial time algorithm, that is based on the observation that all conflicting paths that have the same source and same sink, but do not include the new edge, must be equal because of path independence in the pre-existing graph. Therefore, it suffices to check for one path for a given source and sink, and it follows that all the rest will check out.

In terms of flip points, this is a special instance of case 1, where l and m are the identity.
\begin{center}
\begin{tikzpicture}[scale=0.2]
\tikzstyle{every node}+=[inner sep=0pt]
\draw [black] (25,-31.9) circle (3);
\draw (25,-31.9) node {$S$};
\draw [black] (48,-31.9) circle (3);
\draw (48,-31.9) node {$T$};
\draw [black] (28.9,-19.5) circle (3);
\draw (28.9,-19.5) node {$F_1$};
\draw [black] (43.9,-19.5) circle (3);
\draw (43.9,-19.5) node {$F_2$};
\draw [black] (30.687,-17.11) arc (132.96162:47.03838:8.383);
\fill [black] (42.11,-17.11) -- (41.87,-16.2) -- (41.19,-16.93);
\draw (36.4,-14.36) node [above] {$g_1$};
\draw [black] (41.409,-21.156) arc (-63.93383:-116.06617:11.399);
\fill [black] (41.41,-21.16) -- (40.47,-21.06) -- (40.91,-21.96);
\draw (36.4,-22.82) node [below] {$g_2$};
\draw [black] (23.657,-29.237) arc (-164.08097:-230.83744:7.929);
\fill [black] (23.66,-29.24) -- (23.92,-28.33) -- (22.96,-28.61);
\draw (22.95,-24.04) node [left] {$f_1$};
\draw [black] (30.007,-22.272) arc (11.75211:-46.67052:8.58);
\fill [black] (27.49,-30.26) -- (28.42,-30.08) -- (27.73,-29.35);
\draw (30.56,-27.24) node [right] {$f_2$};
\draw [black] (45.566,-30.169) arc (-134.8387:-188.56886:9.123);
\fill [black] (42.98,-22.34) -- (42.36,-23.06) -- (43.35,-23.21);
\draw (42.57,-27.26) node [left] {$h_2$};
\draw [black] (46.175,-21.439) arc (41.169:-4.57656:10.227);
\fill [black] (46.18,-21.44) -- (46.33,-22.37) -- (47.08,-21.71);
\draw (48.95,-24.27) node [right] {$h_1$};
\end{tikzpicture}
\end{center}

@@

In an N node graph, there are $^NP_2 = N(N-1)$ choices of flip points. Thus we can place a polynomial upper bound on the number of paths that actually need to be verified.

@@

\subsection{Optimized Path Finding Algorithm}
We utilize this idea of selecting flip points to devise a polynomial time path search. This search does not return all flip tolerant paths, but rather, a path for every feasible choice of flipping point.

\begin{verbatim}
let (s,t) be the new edge.
The graph being used for path search does not contain (s,t).

# set of tuples storing conflicting pairs of paths
set_of_pairs = {}

for every node f_1 in the graph:
    for every node f_2 in the graph:
        try:
            find a path p_1 from f_1 to s.
            find a path p_2 from t to f_2.
            let path p= p_1 + (s,t) + p_2.
            find a path q from f_1 to f_2.
            add (p,q) to set_of_pairs    
        catch(path search failed):
            print('There is no conflicting pair from f_1 and f_2 such that
            one of the paths contains the new edge')

equality_check_verify(set_of_pairs)

\end{verbatim}

As observed earlier, the number of paths that this algorithm returns is asymptotically tight.

\subsection{Proof of correctness}

*TODO fill out*

- Ignoring cycle containing paths is ok

- No point in doing a pair unless exactly one has new edge

- Excepting the cases already dismissed, there is no conflicting pair such that its information is not covered by one of the paths in the final set (best way to write might be by cases).
